### 2020-03-10 listofchanges1.tex
1st rev: add motivation for pesq & snrseg

\section*{List of changes}

\newcommand{\pagelineref}[1]{Page~\pageref{#1}, Line~\lineref{#1}}

\begin{itemize}
	\item \pagelineref{ln:daily-tasks}: Replaced
			\begin{quotation}
				For standard \textbf{consumer-grade and even commercial-grade applications}, this tried-and-true method of sampling and systematically discarding the imperceptible information \textbf{works without a hitch}.
			\end{quotation}
		\noindent with
			\begin{quotation}
				For standard \textbf{daily tasks}, this tried-and-true method of sampling and systematically discarding the imperceptible information \textbf{is acceptable for most applications}.
			\end{quotation}
		\noindent The reason for the first correction is that the recommended correction repeated the word ``application'' from the first.
	\item \pagelineref{ln:from-this}: Split long sentence into two shorter sentences; fixed verb tenses.
	\item \pagelineref{ln:mathew}: Added authors' surnames to inline citation.
	\item \pagelineref{ln:theirs}: Split another long sentence.
	\item \pagelineref{ln:motivation} Moved the first sentence (``This metric is a full reference...'') to Subsec.~\ref{ssec:eval} on \pagelineref{ln:pesq-reference}; added the motivation for using PESQ.
	\item \pagelineref{ln:motivation-snrseg}: Added short description and motivation for \snrseg; added the goal of observing the effect of compression ratio/number of sampling windows vs metrics.
\end{itemize}