### 2020-03-12 listofchanges1.tex
2nd rev: rephrase PESQ findings, enlarge figure labels

\section*{List of changes}

\newcommand{\pagelineref}[1]{Page~\pageref{#1}, Line~\lineref{#1}}

\begin{itemize}
	\item \pagelineref{ln:pesq-abstract}: Restated PESQ findings.
	\item \pagelineref{ln:pesq-conc}: Edited conclusion to reflect changes in abstract.
	\item Page~\pageref{fig:pesq-snr}: Made Fig.~2 labels larger.
	\item Removed some extra information and references so as not to exceed 4 pages (pero mukhang mag-eexceed pa rin po).
	\item Abbreviated ``compression ratio'' to ``CR''
\end{itemize}