### 2020-07-03 List of Changes.tex

\documentclass[a4paper,12pt]{article}
\usepackage[margin=1in]{geometry}
\parindent=0pt
\parskip=10pt

\begin{document}

\section*{List of Changes}

\begin{itemize}
	\item \textbf{Abstract}: Incorporated changes suggested by the reviewer.
		\begin{itemize}
			\item Last sentence: removed ``It is shown''
		\end{itemize}

	\item \textbf{Introduction}: Incorporated changes suggested by the reviewer.
		\begin{itemize}
			\item Paragraph 1: removed Eq.~(1) and replaced preceding sentence with ``...sampled at a rate at least twice the signal's highest frequency component.''
			\item Paragraph 2: replaced ``impossible'' with ``infeasible''
		\end{itemize}
	\item \textbf{Methodology}: Incorporated changes suggested by the reviewer.
		\begin{itemize}
			\item Sec.~2.2: replaced ``some domain'' with ``a domain''
		\end{itemize}
	\item \textbf{Results \& Discussion}: Incorporated changes suggested by the reviewer.
		\begin{itemize}
			\item Paragraph 1: replaced ``appears to have decreased'' with ``decreased''
			\item Last sentence: removed ``However''
		\end{itemize}
\end{itemize}

\end{document}